\documentclass[11pt]{article}

\usepackage[margin=1.1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue!60!black,citecolor=blue!60!black,urlcolor=blue!60!black]{hyperref}
\usepackage{microtype}

\newcommand{\br}[1]{\langle #1 \rangle}
\newcommand{\N}{\mathcal{N}}

\title{Executable Positive Geometry:\\
An End-to-End Verification Tour from Bell Tests to the Two-Loop
Amplituhedron, with Holographic Wedge Localization Measured on
Superconducting Hardware}

\author{Jon Torrez\thanks{Correspondence: \texttt{epg@subnet345.com}.
Prepared with Ariadne Vale, an AI research assistant (Claude, Anthropic);
see Acknowledgments.}}

\date{July 5, 2026}

\begin{document}
\maketitle

\begin{abstract}
We report a systematic, fully executable verification tour of the modern
``physics from geometry'' program, spanning three of its pillars:
(i)~quantum nonlocality and error correction, (ii)~emergent geometry from
entanglement, and (iii)~positive geometries underlying scattering
amplitudes. Every result is either derived symbolically in exact
arithmetic, computed numerically without approximation, or measured on IBM
superconducting hardware, with all code and job identifiers retained. On
the amplitudes side we verify, in exact rational/symbolic arithmetic: the
canonical form of the ABHY associahedron against the bi-adjoint $\phi^3$
amplitude; the six-term NMHV identity of the tree amplituhedron
$\mathcal{A}(6,1,4)$ together with its facet structure (Gale evenness,
9~facets); the equality of the one-loop four-point box integrand with a
pure dlog form; and the ``maximally positive numerator'' structure of the
two-loop four-point integrand, whose double-box representation reproduces
the positive part of the mutual condition $\br{ABCD}$ exactly. On the
emergence side we locate the transverse-field Ising transition
variationally, extract the Ising central charge from entanglement
($c \approx 0.58$ vs.\ the exact $1/2$ at $N{=}12$), and construct a
six-tile holographic (HaPPY) network in which reconstruction depth,
quantized information transfer, and a discrete Ryu--Takayanagi law are
verified exactly. On hardware we measure: a CHSH violation
$S = 2.7579$ with a no-signaling check; entanglement-wedge witnesses for a
single holographic tile at $96\%$ of ideal contrast after zero-noise
extrapolation; and, for the 21-qubit six-tile network, bulk-operator
localization---the same boundary avatar of a bulk operator reads
$+0.55$ on the correct tile and $+0.004$ on the wrong one, a
$\sim$130:1 contrast. We also document an instructive failure: zero-noise
extrapolation diverges when circuit depth pushes raw signals to the noise
floor, and the experiment's null witnesses correctly flagged the invalid
values. This paper is a pedagogical verification study rather than a
report of new theory; its contribution is a complete, reproducible chain
from textbook nonlocality to research-frontier positive geometry, with
each link executed rather than cited.
\end{abstract}

\section{Introduction}

Two research programs, developed largely independently, suggest that the
basic furniture of physics---spacetime locality on one side, scattering
processes on the other---may be derived rather than fundamental. In the
holographic program, bulk spacetime emerges from patterns of entanglement
among boundary degrees of freedom \citep{Maldacena:1997re,Ryu:2006bv,
VanRaamsdonk:2010pw,Swingle:2009bg}, with quantum error correction
supplying the mechanism by which bulk information is redundantly and
robustly encoded \citep{Almheiri:2014lwa,Pastawski:2015qua}. In the
amplitudes program, scattering amplitudes of specific quantum field
theories are recovered as canonical differential forms of \emph{positive
geometries} \citep{ArkaniHamed:2017tmz}---the amplituhedron for planar
$\N=4$ super Yang--Mills \citep{ArkaniHamed:2013jha,ArkaniHamed:2013kca},
the ABHY associahedron for bi-adjoint $\phi^3$ theory
\citep{ArkaniHamed:2017mur}, and cosmological polytopes for wavefunction
coefficients \citep{ArkaniHamed:2017fdk}---so that locality and unitarity
appear as consequences of positivity rather than axioms.

Both programs are unusually amenable to direct verification: their core
claims are statements about finite-dimensional linear algebra, polytopes,
rational functions, and few-qubit quantum states. This paper documents a
complete execution of that observation. Rather than survey the
literature, we \emph{run} it: each foundational claim is reproduced as an
exact computation (symbolic or rational arithmetic, deterministic
statevector simulation) or as a measurement on IBM Quantum hardware, and
the failures and corrections encountered along the way are reported
alongside the successes.

We emphasize scope honestly at the outset. Nothing here is new theory.
The hardware experiments are small-scale demonstrations, not competitive
benchmarks. The value of the exercise, we believe, is threefold: it
provides a reproducible pedagogical chain from elementary quantum
information to research-frontier positive geometry; it demonstrates that
several claims usually taken on authority (triangulation independence,
the maximally positive two-loop numerator, quantized wedge information)
can be checked end-to-end on commodity resources; and it supplies a
case study in verification discipline, including one instructive
mitigation failure caught by designed-in null tests.

\section{Nonlocality without signaling}
\label{sec:chsh}

We prepare Bell pairs $\left|\Phi^+\right\rangle$ and evaluate the CHSH
statistic $S = E(a,b) - E(a,b') + E(a',b) + E(a',b')$ with measurement
angles $(0, \pi/2)$ for one party and $(\pi/4, 3\pi/4)$ for the other.
Local realism bounds $S \le 2$; quantum mechanics attains Tsirelson's
bound $2\sqrt{2} \approx 2.8284$.

On a noiseless simulator ($2\times 10^4$ shots per setting) we obtain
$S = 2.8375 \pm 0.01$, consistent with Tsirelson's bound. On the
156-qubit IBM Heron processor \texttt{ibm\_kingston} (job
\texttt{d94sl25gc6cc73ffgo5g}), the four correlators measure
$+0.6968$, $-0.6681$, $+0.6968$, $+0.6962$ against ideal $\pm 0.7071$,
giving
\begin{equation}
S_{\text{hw}} = 2.7579 \pm 0.010,
\end{equation}
violating the classical bound by roughly $75$ standard deviations. In
the same runs, each party's marginal outcome distribution is independent
of the remote setting to within $2\times10^{-4}$: the correlations are
real and yet carry zero signaling capacity. This pair of facts---strong
correlation, no channel---recurs as the organizing constraint of
everything that follows.

\section{Quantum error correction: from repetition to topology}
\label{sec:qec}

\paragraph{Repetition code.} The three-qubit bit-flip code encodes
$\alpha\left|0\right\rangle + \beta\left|1\right\rangle \mapsto
\alpha\left|000\right\rangle + \beta\left|111\right\rangle$ and measures
the parities $Z_0Z_1$, $Z_1Z_2$ via ancillas. Because both codewords have
identical parities, syndrome extraction acquires the \emph{location} of a
bit-flip while learning nothing about the encoded amplitudes; we verify
that an encoded state with $P(1) = 0.750$ survives deterministic
single-qubit attacks with corrected readout $P(1) = 0.746$--$0.753$
across all error positions. Under symmetric bit-flip noise of rate $p$
per qubit, the measured logical error rate tracks the theoretical
$3p^2 - 2p^3$ across $p \in [0.01, 0.55]$, crossing the break-even line
at $p = 1/2$ exactly as predicted: redundancy helps only below threshold.

\paragraph{Surface code.} We implement the distance-3 rotated surface
code (``surface-17'': 9 data, 8 ancilla qubits) with full two-round
quantum syndrome extraction. Deterministic attack fingerprints are
reproduced at 100\% of shots: single $X$, $Z$, and $Y$ errors light
exactly their adjacent checks; the two-qubit string $X_2X_5$ lights only
its \emph{endpoints}; and the logical string $X_6X_7X_8$ produces a
silent syndrome while flipping the logical qubit---the defining
topological statement that the only invisible error spans the lattice.
Under code-capacity bit-flip noise with a brute-force minimum-weight
decoder, the logical error rate scales as $p_L \approx (13$--$15)\,p^2$
at small $p$, confirming that distance 3 removes all single faults.

\paragraph{The fair fight.} Against symmetric $X$ and $Z$ noise
(each at rate $p$, arbitrary stored state; $2\times10^5$ Monte Carlo
trials per point, justified by Gottesman--Knill), the hierarchy inverts
instructively: the repetition code is \emph{worse than no encoding}
(e.g.\ $0.0302$ vs.\ $0.0200$ at $p = 0.01$) because phase errors pass
through it unseen, while surface-17 wins decisively below its
$\sim\!7\%$ pseudo-threshold ($0.0036$ at $p=0.01$) and loses above it
($0.2256$ vs.\ bare $0.1907$ at $p = 0.10$). Full-spectrum protection at
a constant-factor premium, plus a threshold, is the design pattern
inherited by everything holographic below.

\section{Emergent geometry from entanglement}
\label{sec:emergence}

We study the transverse-field Ising chain
$H = -J\sum_i Z_i Z_{i+1} - h \sum_i X_i$.

\paragraph{Variational location of the transition.} A 48-parameter
hardware-efficient ansatz optimized by COBYLA reproduces exact
ground-state energies for $N = 8$ across $h/J \in [0.2, 2.0]$, with the
order parameter $\br{Z_iZ_{i+1}}$ collapsing from $0.994$ to $0.228$
through the critical region. The variational error peaks at criticality
($0.26$ at $h = J$ vs.\ $0.03$ deep in the ordered phase)---itself a
diagnostic: critical states are the most entangled and hence hardest for
shallow circuits.

\paragraph{Critical entanglement.} For $N = 12$ (exact
diagonalization), the half-chain entropy exhibits the well-known
subtlety that the ordered phase contributes a constant $\ln 2$ of
cat-state (GHZ) entropy. An infinitesimal symmetry-breaking field
($h_z = 10^{-3}$) collapses the cat and exposes the genuine critical
peak ($S = 0.567$ near $h \approx 0.8$ at this size, vanishing in both
phases). At $h = J$, fitting the Calabrese--Cardy law
$S(\ell) = \tfrac{c}{6}\ln\!\big[\tfrac{2N}{\pi}\sin\tfrac{\pi\ell}{N}\big]
+ \text{const}$ yields $c = 0.58$, within finite-size expectations of
the exact Ising value $c = 1/2$.

\paragraph{Entanglement as a ruler.} Defining
$d(i,j) = -\ln I(i,j)$ from two-site mutual information, the disordered
phase ($h = 2J$) yields distances growing linearly with separation
(1.6 per site: an ordinary ruler with a correlation length), while the
critical chain yields logarithmic growth (scale-free connectivity). One
coupling constant selects between two emergent geometries---the
one-dimensional shadow of the tensor-network/holography correspondence
\citep{Swingle:2009bg,Czech:2015qta}.

\section{Positive geometry and scattering amplitudes}
\label{sec:positive}

All results in this section are exact (symbolic or rational arithmetic).

\paragraph{The positive Grassmannian.} For $\mathrm{Gr}(2,4)$ we verify
the Pl\"ucker relation symbolically for a generic matrix, measure that
only $2.14\%$ of Gaussian-random 2-planes are totally positive, and
confirm that the moment-curve parametrization renders every minor
$t_j - t_i > 0$ automatically: positivity is equivalent to cyclic
ordering. In a gauge fixing columns 1 and 3, total positivity reduces to
a positive orthant with $p_{24} = ad + bc$ implied, and the cyclic
product $p_{12}p_{23}p_{34}p_{14} = abcd$: the Parke--Taylor denominator
\emph{is} the canonical form of $\mathrm{Gr}_{+}(2,4)$.

\paragraph{Canonical forms and triangulation independence.} We verify
the recursive definition of canonical forms
\citep{ArkaniHamed:2017tmz} bottom-up: interval residues $\pm 1$;
triangle edge-residues reproducing interval forms; and the unit square
triangulated across both diagonals, with the spurious diagonal pole
cancelling exactly in each sum and both triangulations agreeing.

\paragraph{The ABHY associahedron.} Imposing only masslessness sum
rules and the ABHY constants \citep{ArkaniHamed:2017mur}, a symbolic
solve produces the $n = 5$ pentagon facets
$\{x,\, y,\, c_{13} - x + y,\, c_{13} + c_{14} - x,\,
c_{14} + c_{24} - y\}$. The five-term vertex sum (one planar Feynman
diagram per vertex) combines over a common denominator equal exactly to
the product of the five facets; residues on facets reproduce the
boundary interval forms; vertex counts follow the Catalan numbers.
Factorization---normally a theorem requiring locality and unitarity---is
here the tautology that a boundary of a face is a face.

\paragraph{The tree amplituhedron $\mathcal{A}(6,1,4)$.} With momentum
twistors on a moment curve and $Y$ interior, the two Britto--Cachazo--Feng--Witten (BCFW)
triangulations of the NMHV hexagon polytope,
$R_1 = [12345]+[12356]+[13456]$ and $R_2 = [23456]+[12456]+[12346]$,
agree \emph{exactly} in rational arithmetic ($R_1 - R_2 = 0$ at three
random interior points)---the six-term NMHV identity as a statement
about triangulations. Approaching the spurious hyperplane
$\br{Y1235} \to 0$, individual terms diverge as $\pm 86.8$ while the sum
holds at $1.5\times10^{-4}$. An exact same-side-of-hyperplane test
classifies all $\binom{6}{4} = 15$ vertex 4-subsets into 9 facets---
precisely the Gale-evenness set $\{i, i{+}1, j, j{+}1\}$, i.e.\ the
physical propagator poles---and 6 interior seams.

\paragraph{One loop.} In the single positive cell of the $n=4$ MHV loop
amplituhedron \citep{ArkaniHamed:2013kca,Kojima:2020gxs}, parametrized
by $A = Z_1 + xZ_2 + wZ_3$, $B = -Z_1 + yZ_3 + zZ_4$, all physical
brackets are monomials ($\br{AB12} = wz$, $\br{AB23} = z$,
$\br{AB34} = x$, $\br{AB14} = xy$), the sign-flip sequence is
$\{+,-,+\}$, and the box integrand divided by
$d\log x \wedge d\log y \wedge d\log z \wedge d\log w$ equals $1$
identically.

\paragraph{Two loops.} For two loop lines the mutual condition
evaluates to
\begin{equation}
\br{ABCD} = (z_1 - z_2)(w_1 x_2 - w_2 x_1) + (x_1 - x_2)(y_1 z_2 - y_2 z_1),
\end{equation}
and Monte Carlo sampling shows only $49.8\%$ of independently valid loop
pairs satisfy it: the geometry entangles the loops. Converting the
physical double-box representation (s- and t-channel plus loop
relabelings) to this chart and multiplying by $\br{ABCD}$ yields the
numerator
\begin{equation}
N = w_1 x_2 z_1 + w_2 x_1 z_2 + x_1 y_1 z_2 + x_2 y_2 z_1,
\end{equation}
which is \emph{exactly} the sum of the positive monomials of
$\br{ABCD}$---the ``maximally positive numerator'' phenomenon
\citep{Kojima:2020gxs}, here recovered from raw field theory. Two errors
were caught by this cross-check during the work: an incorrect channel
assignment in the double-box term list (diagnosed by a numerator with
missing monomials) and the ordered-pair symmetrization convention
(diagnosed by an overall factor of 2). We record these deliberately: the
positivity structure functioned as a debugger for the physics input.

\section{Holographic codes: exact results and hardware measurements}
\label{sec:happy}

\subsection{A single tile, exactly}

The six-leg perfect tensor of the HaPPY code
\citep{Pastawski:2015qua}---the encoding isometry of the
$[[5,1,3]]$ code, with the bulk leg entangled to a reference
$R$---satisfies, in exact statevector arithmetic: all 20 three-leg
bipartitions maximally entangled ($S = 3\ln 2$); mutual information
$I(R{:}A) = 0$ for every boundary region with $|A| \le 2$ and
$I(R{:}A) = 2\ln 2$ for every $|A| \ge 3$ (the sharp entanglement-wedge
transition); and weight-3 boundary representatives of the logical
operators obtained by stabilizer multiplication,
$\bar X = -(IYYIX)$ on qubits $\{2,3,5\}$ and $-(XIYYI)$ on
$\{1,3,4\}$, verified to act identically on the bulk
(operator pushing; cf.\ \citep{Almheiri:2014lwa}).

\subsection{A single tile on hardware}

We prepared the tile state on \texttt{ibm\_kingston} via a synthesized
Clifford circuit (42 two-qubit gates after routing) and measured Pauli
witnesses (Table~\ref{tab:tile}): stabilizers, weight-6 logical
correlators, three wedge avatars, and four null correlators between $R$
and two-qubit regions. Raw (readout-mitigated) values average $+0.535$
for wedge witnesses. Adding zero-noise extrapolation (noise factors
$1,3,5$; exponential fit), Pauli twirling, and dynamical decoupling
raises the wedge mean to $+0.9558$ while the null mean \emph{decreases}
to $0.0046$; two witnesses slightly exceed $+1$, the expected
variance signature of an unbiased extrapolation. The asymmetry is the
point: mitigation restores contrast that noise removed but cannot
create correlations forbidden by the code---minority regions stay dark.

\begin{table}[t]
\centering
\small
\begin{tabular}{lrrr}
\toprule
Witness & Ideal & Raw & Mitigated \\
\midrule
Stabilizer $S_1$ & $+1$ & $+0.9300$ & $+0.9949$ \\
Stabilizer $S_2$ & $+1$ & $+0.4235$ & $+0.9996$ \\
Stabilizer $S_3$ & $+1$ & $+0.4466$ & $+0.9003$ \\
Stabilizer $S_4$ & $+1$ & $+0.4221$ & $+0.8897$ \\
$X_R \bar X$ (weight 6) & $+1$ & $+0.4450$ & $+1.0504$ \\
$Z_R \bar Z$ (weight 6) & $+1$ & $+0.4036$ & $+0.9668$ \\
Bulk-$X$, wedge $\{2,3,5\}$ & $+1$ & $+0.4351$ & $+1.0343$ \\
Bulk-$X$, wedge $\{1,3,4\}$ & $+1$ & $+0.7707$ & $+0.9411$ \\
Bulk-$Z$, wedge $\{1,4,5\}$ & $+1$ & $+0.3993$ & $+0.8921$ \\
Null $X_R X_1 X_2$ & $0$ & $+0.0015$ & $-0.0021$ \\
Null $Z_R Z_4 Z_5$ & $0$ & $-0.0146$ & $-0.0146$ \\
Null $X_R Z_1 Y_2$ & $0$ & $+0.0227$ & $+0.0000$ \\
Null $Z_R X_2 X_3$ & $0$ & $-0.0215$ & $+0.0017$ \\
\midrule
Mean wedge & $+1$ & $+0.5350$ & $+0.9558$ \\
Mean $|$null$|$ & $0$ & $0.0151$ & $0.0046$ \\
\bottomrule
\end{tabular}
\caption{Single-tile witnesses on \texttt{ibm\_kingston}
($10^4$ shots; raw job \texttt{d94tqhlgc6cc73ffi0eg}, mitigated job
\texttt{d94ttgtgc6cc73ffi3p0}).}
\label{tab:tile}
\end{table}

\subsection{The six-tile network, exactly}

Gluing five petal tiles to a central tile (6 bulk qubits, 5 bonds, 20
boundary qubits) yields, in exact contraction: reconstruction depth
(a petal bulk qubit is recovered by a 3-qubit boundary arc; the central
qubit requires 10 of 20); \emph{quantized} information transfer, with
$I(R{:}\text{arc})$ taking only the values $\{0, 1, 2\} \times \ln 2$
(the intermediate shelf corresponding to a single logical Pauli inside
the wedge); and a discrete Ryu--Takayanagi law, the arc entropies
$S/\ln 2 = 1,2,2,1,2,3,3,2,3,3$ for lengths $1$--$10$ tracking the
minimal bond cut, including the characteristic dips when the cut snaps
from boundary legs to an internal bond.

\subsection{The six-tile network on hardware}

The 21-qubit network state is a stabilizer state; we derived its 21
generators numerically by pushing input Paulis through the petal
isometries (choosing minimum-weight representatives from the Pauli
decomposition of $W\sigma W^\dagger$) and synthesized a preparation
circuit (148 two-qubit gates; 553 after routing) whose statevector
matches the exact tensor-network state to overlap $1.0000000000$.

A first hardware attempt with zero-noise extrapolation (job
\texttt{d94u25nu62ks7396fhlg}) \emph{diverged}: null witnesses returned
values of order $\pm 4\times10^{9}$, impossible for Pauli expectations.
The cause is generic and worth recording: at $\sim$550 two-qubit gates,
several raw signals sit at the noise floor already at unit noise, and
exponential extrapolation through noise-floor points amplifies fit
noise without bound. The null tests functioned exactly as designed,
flagging the entire dataset---including superficially plausible
entries---as artifacts.

The clean rerun (job \texttt{d94u424ql68s73c9vtig}; twirling, readout
mitigation, dynamical decoupling, no extrapolation) gives physical
values throughout (Table~\ref{tab:flower}). The headline is the
localization contrast:
\begin{equation}
\underbrace{+0.5519}_{\text{bulk-}X\text{ avatar on petal 1}}
\quad\text{vs.}\quad
\underbrace{+0.0042}_{\text{identical operator on petal 2}}
\end{equation}
a $\sim$130:1 ratio at $\sim$55$\sigma$: the same boundary operator
reads the bulk qubit through the correct tile and reads exactly nothing
through the wrong one. The remaining wedge witnesses survive at reduced
contrast ($+0.099$, $+0.065$; both above the null floor), and the
per-petal encoding checks ($0.07$--$0.24$) quantify how badly 553 noisy
gates bruise the state. A bulk point of a six-tile emergent geometry
was located---and localized---on physical hardware, at the edge of what
unmitigated depth allows.

\begin{table}[t]
\centering
\small
\begin{tabular}{lrr}
\toprule
Witness & Ideal & Measured \\
\midrule
Image check, petal 1 & $+1$ & $+0.0884$ \\
Image check, petal 2 & $+1$ & $+0.0692$ \\
Image check, petal 3 & $+1$ & $+0.1218$ \\
Image check, petal 4 & $+1$ & $+0.1979$ \\
Image check, petal 5 & $+1$ & $+0.2424$ \\
Wedge: bulk-$X$ avatar A, petal 1 & $+1$ & $+0.5519$ \\
Wedge: bulk-$X$ avatar B, petal 1 & $+1$ & $+0.0993$ \\
Wedge: bulk-$Z$ avatar, petal 1 & $+1$ & $+0.0648$ \\
Null: same bulk-$X$ avatar, petal 2 & $0$ & $+0.0042$ \\
Null: $X_R XX$, petal-1 pair & $0$ & $+0.0088$ \\
Null: $Z_R ZZ$, petal-3 pair & $0$ & $-0.0109$ \\
\bottomrule
\end{tabular}
\caption{Six-tile flower network on \texttt{ibm\_kingston}
($10^4$ shots, job \texttt{d94u424ql68s73c9vtig}; twirling + readout
mitigation + dynamical decoupling, no extrapolation).}
\label{tab:flower}
\end{table}

\section{Discussion}
\label{sec:discussion}

Three observations organize what the exercise taught us.

\paragraph{One trick, three scales.} The repetition code protects a bit
by majority; the surface code protects a qubit by homology; the
holographic code protects a bulk point by wedge structure. In each case
information is stored in a global property that no sub-threshold local
process reaches, and in each case the protection pattern \emph{is} a
geometry (a line, a torus, a hyperbolic disk). Our hardware results
trace the same arc operationally: near-perfect contrast at 2 qubits,
$96\%$ mitigated contrast at 6, a surviving-but-bruised $55\sigma$
signal at 21. The scaling wall is exponential in circuit depth, and the
known way through it is the error correction of
Section~\ref{sec:qec}---which is also the mechanism under study. The
experiments are, in a precise sense, error-correction-limited studies of
error correction.

\paragraph{Positivity as output specification.} At tree level,
positivity fixed pole locations; at one loop, it reduced the box to a
pure dlog form; at two loops it fixed the \emph{numerator} exactly.
Twice during this work, discrepancies between a hand-supplied physics
formula and the positive-geometry prediction turned out to be errors in
the former. Whatever positive geometry ultimately is, it is at minimum
an unusually effective consistency engine for perturbative field
theory.

\paragraph{No levers.} Every prospective ``exploit'' the underlying
emergent-spacetime narrative might suggest was tested and closed by the
physics itself: entanglement carries no signal
(Section~\ref{sec:chsh}); minority wedges carry no information
(Sections~\ref{sec:happy}); adjacency in kinematic space is descriptive,
with positivity \emph{enforcing} rather than circumventing locality and
causality (Section~\ref{sec:positive}). The constraints are not
guardrails around the structure; they are the structure.

\paragraph{Limitations.} All models studied are toys: the amplituhedron
addresses planar $\N=4$ super Yang--Mills; the associahedron, bi-adjoint
$\phi^3$; the HaPPY network, a static AdS-like geometry. Whether our
universe employs any of these mechanisms is open; de Sitter holography
and the emergence of \emph{time} have no comparable toy models at all.
The hardware demonstrations use few qubits, favorable witnesses, and a
free-tier device; they demonstrate structure, not capability. The
recent two-loop geometry literature we verified against
\citep{Dian:2024tvi} is young and single-group.

\section{Methods}

All computations used Python 3.13 (NumPy/SciPy/SymPy) and Qiskit 2.5
with Aer for exact simulation; hardware runs used IBM Quantum
(\texttt{ibm\_kingston}, Heron, 156 qubits) via the EstimatorV2/SamplerV2
primitives with $10^4$ shots per observable, and (where stated)
resilience options: twirled readout mitigation, Pauli gate twirling,
XY4 dynamical decoupling, and zero-noise extrapolation with gate
folding. Exact claims use symbolic or rational arithmetic throughout;
no floating-point equality is asserted anywhere in
Section~\ref{sec:positive}. Stabilizer-state preparation circuits were
synthesized from explicit generator lists and verified against exact
statevectors before submission. All literature-derived formulas
(one-loop cell parametrization, two-loop mutual condition and numerator
structure, amplituhedron region definitions) were checked against their
sources \citep{Kojima:2020gxs,Dian:2024tvi,ArkaniHamed:2013kca} rather
than reproduced from memory. Complete source code (thirteen laboratory
scripts), an annotated 18-paper reference corpus with evidence tiers,
and all IBM job identifiers are retained by the author.

\section*{Acknowledgments}

This work was carried out in collaboration with \emph{Ariadne Vale}, an
AI research assistant (Claude, Anthropic), which wrote and executed the
laboratory code, performed the literature verification, and co-drafted
this manuscript under the author's direction. All hardware runs used
the IBM Quantum open access plan. We acknowledge the use of IBM Quantum
services; the views expressed are those of the author and do not
reflect the official policy of IBM.

\bibliographystyle{unsrtnat}
\bibliography{refs}

\end{document}
