\documentclass[11pt]{article}

\usepackage[margin=1.15in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue!60!black,citecolor=blue!60!black,urlcolor=blue!60!black]{hyperref}
\usepackage{microtype}

\newcommand{\ket}[1]{\lvert #1 \rangle}
\newcommand{\bra}[1]{\langle #1 \rvert}

\title{Magic of cosmological particle production and the de Sitter horizon}

\author{Jon Torrez\thanks{Correspondence: \texttt{epg@subnet345.com}.
Computations performed by two AI research agents (Ariadne Vale, a quantum
physics persona, and codex-science), each result independently
cross-checked; all code at
\url{https://github.com/jrtorrez31337/executable-positive-geometry}.}}

\date{July 2026}

\begin{document}
\maketitle

\begin{abstract}
Non-stabilizerness (``magic'') has been proposed as the quantum-information
resource behind gravitational back-reaction, but that program lives almost
entirely in anti--de Sitter and Sachdev--Ye--Kitaev (SYK) settings. We ask
whether magic is computable, and behaves, in two settings closer to our own
universe: cosmological particle production and the de Sitter horizon.
Two results, both in deliberately conservative toy models. (i) Cosmological
pair creation sources magic \emph{by resource theory}: a free bosonic
two-mode-squeezed out-state is Gaussian and hence Wigner-positive
(magic-free), with any finite-qubit stabilizer entropy an encoding artifact;
only genuine non-Gaussianity (interactions) turns on Wigner negativity. The
fermionic out-state, by contrast, has exact qubit non-local magic
$-\log_2(4P^2-6P+3)$ that is non-monotonic in the occupation, and on the
sudden de Sitter$\to$radiation transition this fermionic magic is strongly
suppressed unless the transition is violent---so cosmological fermion magic
is a probe of the violence of reheating. (ii) The de Sitter horizon, via its
conjectured double-scaled-SYK dual, shows \emph{no} magic signature at the
static level, but a robust one in the dynamics: the stabilizer entropy of
the time-evolved thermofield double grows about twice as fast for a chaotic
Hamiltonian as for a free one, and the chaotic--free late-time gap grows
with system size ($+0.61,+1.03,+1.37$ at $N=8,10,12$; positive under a
same-spectrum control for $N\ge 10$), the de Sitter analog of the
magic$\leftrightarrow$area link. We state throughout what is a theorem, what
is a conservative proxy, and where an initially over-stated claim was
narrowed by adversarial cross-check.
\end{abstract}

\section{Introduction}

Entanglement entropy fixes the Ryu--Takayanagi \emph{area}, but an area on a
fixed background does not gravitate; producing state-dependent geometry and
back-reaction requires a further resource, and recent work identifies it with
non-stabilizerness, or ``magic''
\citep{cao_area,cao_backreaction,cao_codes}. That identification has been
developed in AdS holographic codes and in SYK. Here we ask the two natural
``our universe'' questions it invites: does cosmological particle production
generate magic (Section~\ref{sec:p1}), and is the de Sitter horizon magical
(Section~\ref{sec:p2})?

Both are addressed by exact computation in small toy models, reusing a
Fast-Walsh--Hadamard stabilizer-entropy engine and standard SYK machinery.
The organizing principle that emerges is a \emph{resource-theory} one: in the
continuous-variable description magic is Wigner negativity, and free
(Gaussian) sectors are magic-free by Hudson's theorem, so any magic must come
from either the genuinely discrete (fermionic) sector or from interactions
and chaos.

\section{Magic of the expanding vacuum}
\label{sec:p1}

An expanding background turns the in-vacuum into an out-state of created
pairs of opposite comoving momentum---a two-mode-squeezed (boson) or single
Pauli-blocked pair (fermion) per mode. We compute the magic of that
out-state.

\subsection{Boson: a theorem-anchored null, and the non-Gaussian turn-on}
For a free scalar mode pair, cosmological particle production gives the
two-mode squeezed vacuum
\begin{equation}
  \ket{\mathrm{TMSV}(r)}
  = \sqrt{1-\lambda^2}\sum_{n=0}^{\infty}\lambda^n
    \ket{n_k,n_{-k}},\qquad \lambda=\tanh r .
\end{equation}
This is a pure Gaussian continuous-variable state. By Hudson's theorem, and
equivalently by the continuous-variable Gottesman--Knill simulation criterion
\citep{hudson,mari_eisert}, its Wigner function is nonnegative for every
squeezing parameter $r$. Its physical CV mana,
\begin{equation}
  \mathcal{M}_{\rm CV}=\log_2\!\int dq\,dp\, |W(q,p)| ,
\end{equation}
is therefore exactly zero. Free bosonic inflation can create large
entanglement and large occupation, but not Wigner negativity; this agrees with
the recent ``Inflation is Not Magic'' analysis \citep{inflation_not_magic}.

The important guardrail is that a finite Fock cutoff followed by a binary
encoding is not the same resource theory. Truncating
$\ket{\mathrm{TMSV}}$ to $2^q$ oscillator levels and mapping the Fock labels
to qubits can produce a nonzero qubit Pauli stabilizer entropy, but that
number depends on the cutoff and on the arbitrary relabeling of Fock states
once the encoding is large enough to admit non-affine permutations. The same
underlying CV state still has $\mathcal{M}_{\rm CV}=0$. Thus the free-boson
result is not a small qubit signal; it is a theorem-level null in the
invariant CV phase-space resource theory.

Interactions supply the missing resource. As a minimal executable proxy for
primordial non-Gaussianity generated by interactions, we used a cubic-phase
deformation of a Gaussian wavepacket,
\begin{equation}
  \begin{aligned}
  \psi_{\sigma,\gamma}(x)
  &= \sigma^{-1/2}\pi^{-1/4} \\
  &\quad{}\times
    \exp\!\left[-\frac{1}{2}(x/\sigma)^2
      +\frac{i\gamma}{3}(x/\sigma)^3\right].
  \end{aligned}
\end{equation}
At $\gamma=0$ the direct phase-space quadrature gives zero negative volume.
For $\gamma>0$, Wigner-negative interference fringes appear: in our grid
calculation the negative volume grows from $6.9\times10^{-4}$ at
$\gamma=0.24$ to $7.4\times10^{-3}$ at $\gamma=0.40$ and
$3.3\times10^{-2}$ at $\gamma=0.70$. A toy ultra-slow-roll proxy
$\gamma_{\rm eff}=\gamma_0 e^{2\Delta N}$ reproduces the expected rapid
growth once the non-Gaussianity becomes appreciable, matching the qualitative
mechanism of \citet{ireland_vennin}. The claim is deliberately narrow: this
is not a reproduction of their full inflationary EFT calculation, but it
isolates the resource-theory transition. Linear squeezing alone remains
Gaussian and Wigner-positive; genuine non-Gaussianity is what turns bosonic
cosmological magic on.

\subsection{Fermion: exact qubit magic and the violence of reheating}
Pauli exclusion makes the fermionic out-state a single pair per mode. Under
the Jordan--Wigner map---which is exact because fermionic Fock space is
two-dimensional per mode---mode $k$ and its partner form the two-qubit state
$\alpha\ket{00}+\beta\ket{11}$ with $|\beta|^2$ the occupation. Its qubit
non-local magic (the stabilizer Rényi entropy minimized over local
unitaries) has the closed form
\begin{equation}
  \mathcal{M}(\beta) \;=\; -\log_2\!\big(4P^2-6P+3\big),
  \qquad P=|\alpha|^4+|\beta|^4,
\end{equation}
verified to machine precision against a direct local-unitary minimization.
It is \emph{non-monotonic}: it vanishes at the product limit
($|\beta|^2=0$) \emph{and} at the maximally-mixed ``Bell'' point
($|\beta|^2=\tfrac12$), and peaks at $\mathcal{M}=\log_2\tfrac43\approx0.415$
where $|\beta|^2=\tfrac12(1\mp1/\sqrt2)$. Cosmological creation thus writes
magic into an intermediate band of the occupation.

The physically sharp statement rides the sudden de Sitter$\to$radiation
transition whose scalar Bogoliubov coefficients are analytic
\citep{byun_dsrad}, $|\beta_k|^2=1/(4x^4)$ with $x=k|\eta_e|$. The boson
carries \emph{no} magic across this whole spectrum (Gaussian, above). A
massless fermion is conformally invariant in a conformally-flat FRW universe
and is not produced at all---an exact null. A massive fermion is produced,
but for the paper's smooth ($C^1$) transition the production is strongly
suppressed (peak occupation $\sim3\times10^{-4}$): the mass term $M=m\,a(\eta)$
is $C^1$ and never couples to the $a''/a$ jump that drives boson production.
Only a genuinely violent (sudden, $a'$-discontinuous) transition, for which
the sudden-approximation occupation
$n_k=\tfrac12\big(1-x/\sqrt{x^2+\mu^2}\big)$ runs from $\tfrac12$ in the
infrared to $0$ in the ultraviolet, drives the occupation through the magic
band. Hence
\begin{center}
\emph{cosmological fermion magic is a probe of the violence of reheating:}
\end{center}
negligible for a smooth transition, band-level only for a sudden one, and
zero for a massless (conformal) fermion. The boson/fermion asymmetry has a
one-line root: fermionic Fock space is two-dimensional per mode, so its
qubit magic is physical, whereas bosonic Fock space is infinite-dimensional
and any qubit magic is a truncation artifact.

\section{Is the de Sitter horizon magical?}
\label{sec:p2}

The de Sitter static patch is conjecturally dual to double-scaled SYK
\citep{dsjt_dssyk,doubled_dssyk}, giving a computable stand-in for the
chaotic finite horizon Hilbert space; the honest dictionary is
high-temperature and degree-of-freedom based, so we test scaling with a
horizon-dof proxy rather than a temperature monotone. As a conservative
executable proxy we use ordinary SYK$_4$ (chaotic) versus SYK$_2$ (free).

\paragraph{The static level is a null.} At a fixed time, magic does not
separate chaotic from free: the thermal-pure-quantum stabilizer entropy is
extensive but barely differs between SYK$_4$ and SYK$_2$; the twin-controlled
thermal-state magic even runs backwards (chaos flattens the spectrum toward
maximal mixedness); and the infinite-temperature doubled thermofield double
$\sum_E\ket{E}_L\ket{E^*}_R=\sum_i\ket{i}_L\ket{i}_R$ is exactly the Bell
state, magic-free independent of the Hamiltonian. This null is confirmed
across multiple measures.

\paragraph{The dynamics reveals a signature.} The discriminator is the
stabilizer entropy of the time-evolved thermofield double
$\ket{\mathrm{TFD}(\beta,t)}\propto\sum_n e^{-\beta E_n/2}e^{-iE_nt}
\ket{n}_L\ket{n}_R$, whose magic dynamics is expected to track the spectral
form factor \citep{tfd_sff}. Two discriminators, both with error bars over
independent disorder realizations:
\begin{itemize}
\item \textbf{Scrambling rise.} Chaotic SYK$_4$ thermofield-double magic
grows about twice as fast as free SYK$_2$ under evolution; the rise ratio is
$2.05,2.30,2.55,2.17$ at $N=6,8,10,12$---robust across size.
\item \textbf{Extensive plateau gap.} The late-time SYK$_4$--SYK$_2$ magic gap
is positive and grows with the number of degrees of freedom,
$+0.61,+1.03,+1.37$ at $N=8,10,12$ (tight seeds at $N=12$,
$+1.37\pm0.07$). Under a same-spectrum diagonal-basis control that removes the
eigenbasis contribution the gap is $-0.14,+0.75,+0.94$ at $N=8,10,12$:
positive and growing for $N\ge10$.
\end{itemize}
Strikingly, at a fixed time the \emph{free} system carries more magic (the
static gap is negative); only under evolution does the chaotic system reveal
its larger magic capacity. The growth of the gap with the horizon-dof count
is the de Sitter analog of the magic$\leftrightarrow$area link
\citep{cao_area}, seen dynamically where the static snapshot is flat.

\section{Scope, and a note on method}

Every model here is a toy: the fermion result is a two-level Dirac mode
model, and the de Sitter results use a conservative SYK$_4$/SYK$_2$ proxy for
double-scaled SYK at finite $N\le12$, not the full constrained double-scaled
construction. The claims are correspondingly bounded: the boson null is a
theorem, the fermion reheating-violence statement is a mechanism in a toy
transition, and the horizon-magic signature is a scrambling-sensitive
diagnostic whose extensive scaling is firmed for $N\ge10$ but wants larger
$N$ / full DSSYK to become a theorem.

We record one methodological point deliberately. The extensive plateau gap
was first reported from a single disorder seed as a clean trend from small
$N$; an independent-seed error-bar cross-check showed the smallest-$N$ point
was a fluctuation (an actual sign crossover), and the claim was narrowed to
its firmed form above. The correction, and the two agents' mutually
adversarial cross-checks that produced it, are part of the result.

\begin{thebibliography}{99}
\small
\bibitem{cao_area} C.~Cao \emph{et al.}, \emph{Non-trivial Area Operators
Require Non-local Magic}, arXiv:2306.14996.
\bibitem{cao_backreaction} \emph{Gravitational back-reaction is magical},
arXiv:2403.07056.
\bibitem{cao_codes} C.~Cao, W.~Cheng, A.~Karthikeyan, Y.~Li, J.~Preskill,
\emph{State-dependent geometries from magic-enriched quantum codes},
arXiv:2603.13475.
\bibitem{hudson} R.~L.~Hudson, \emph{When is the Wigner quasi-probability
density non-negative?}, Reports on Mathematical Physics 6, 249--252 (1974).
\bibitem{mari_eisert} A.~Mari, J.~Eisert, \emph{Positive Wigner functions
render classical simulation of quantum computation efficient}, PRL 109,
230503 (2012).
\bibitem{ireland_vennin} L.~Ireland, V.~Vennin, arXiv:2601.22219.
\bibitem{inflation_not_magic} S.~S.~Haque, Jafari, B.~Underwood,
\emph{Inflation is Not Magic}, arXiv:2512.10126.
\bibitem{byun_dsrad} Byun, Kim, Lee, arXiv:2605.04099.
\bibitem{dsjt_dssyk} \emph{dS JT Gravity and Double-Scaled SYK},
arXiv:2209.09997.
\bibitem{doubled_dssyk} arXiv:2310.16994.
\bibitem{tfd_sff} \emph{Connecting Magic Dynamics in Thermofield Double
States to Spectral Form Factors}, arXiv:2601.12787.
\end{thebibliography}

\end{document}
