\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{xcolor}
\usepackage[colorlinks=true,linkcolor=blue!60!black,citecolor=blue!60!black,urlcolor=blue!60!black]{hyperref}
\usepackage{braket}

\newcommand{\Sfour}{S_4}
\newcommand{\code}{[[5,1,3]]}

\title{Time reversal is the one twist a self-referential code cannot enact:\\
a logical-holonomy census of the $\code$ code}
\author{Ariadne Vale and Tobin Masala\\[2pt]
\small Executable Positive Geometry \\[-2pt]
\small \texttt{epg@subnet345.com}}
\date{13 July 2026}

\begin{document}
\maketitle

\begin{abstract}
An embedded observer who can only read and re-encode a quantum system --- never
step outside it --- can enact only unitary operations on its own logical data.
We make this concrete for the $\code$ perfect code. A closed loop in
code-deformation space is an automorphism of the stabilizer group realized by a
qubit permutation composed with a uniform transversal single-qubit Clifford; its induced
action on the logical Bloch frame is a \emph{holonomy}. Enumerating all
$120\times24=2880$ candidate loops, the $240$ automorphisms realize exactly $24$
distinct logical holonomies. They form the octahedral group $\Sfour$ --- the full
logical Clifford group modulo Pauli's --- and \emph{every one is a proper
rotation}, $\det R=+1$. The orientation-reversing sector absent from this group
is \emph{antiunitary}; by Wigner's theorem, time reversal is implemented in this
antiunitary sector. Hence the one kind of twist an embedded self-reader cannot
acquire by closed unitary self-transport is running the film backwards. We verify
the nontrivial (unitary) holonomy on
superconducting hardware: a $120^\circ$ logical-frame rotation is measured with
$+0.898$ overlap against the rotated prediction versus $-0.003$ unrotated (IBM
\texttt{ibm\_kingston}, job \texttt{d958plnu62ks7396rgmg}). The exclusion itself
is generic to any unitary framework; our contribution is the concrete,
hardware-backed instantiation that the \emph{reachable} inside-twists form a
definite nontrivial group whose missing sector is the antiunitary, time-reversing
one.
\end{abstract}

\section{Setup: what twists can an inside observer acquire?}

This note isolates one sharp consequence of a recurring theme in the
Executable Positive Geometry program: every structure the program builds is
timeless, spatial relations are freely tradeable, and temporal orientation is
the quantity that resists being manufactured or reversed from within a
self-contained description. Here ``from within'' is made precise as the group of
operations available to an observer who is a subsystem of a quantum
error-correcting code and can only permute and locally re-encode the physical
qubits while keeping the code --- the rulebook of self-description --- invariant.

The $\code$ code encodes one logical qubit in five physical qubits and is the
smallest perfect code~\cite{laflamme}; it is also the elementary tile of the
HaPPY holographic code~\cite{happy}, so its logical data is a toy bulk degree of
freedom reconstructed from a boundary. A \emph{code-deformation loop} is an
operation
\begin{equation}
  W \;=\; P_\sigma\, U^{\otimes 5},
  \qquad U \in \mathrm{Clifford}_1,\ \ \sigma\in S_5,
\end{equation}
a physical-qubit permutation $P_\sigma$ composed with a transversal
single-qubit Clifford applied uniformly to all five code qubits, chosen so that
$W$ maps the stabilizer group of the code
\emph{exactly back to itself}. The code returns home; the logical qubit need
not. The induced logical operation has a Bloch-frame representative
$R(W)\in \mathrm{SO}(3)$: the loop's \textbf{holonomy}, the measurable residue of
transporting self-knowledge around a closed path of self-descriptions. The
Klein-bottle question that motivated the census is sharp: does any closed
unitary loop nevertheless realize the mirror action one would need to reverse
the orientation of the logical frame?

\section{The census}

We enumerate all $|S_5|\times|\mathrm{Clifford}_1| = 120\times 24 = 2880$
candidate loops in this uniform-transversal family, retain those that are genuine stabilizer automorphisms,
compute each one's action on the logical frame, and assemble the holonomy
group. The computation is exact (statevector arithmetic on the codewords built
in \texttt{happy\_tile.py}); it is reproduced by \texttt{holonomy.py} in the
repository.

\paragraph{Result.} Of the $2880$ candidates, $240$ are automorphisms. Their
logical actions collapse to \textbf{$24$ distinct holonomies}, with element
orders
\begin{equation}
  \{\,1{:}1,\ 2{:}9,\ 3{:}8,\ 4{:}6\,\},
\end{equation}
the conjugacy-class profile of the octahedral group $\Sfour$. This is precisely
the logical Clifford group modulo Pauli's, $\mathrm{Clifford}_1/\mathrm{Pauli} \cong
\Sfour$: closed journeys through rulebook-space realize the full single-qubit
Clifford symmetry of the logical frame. Sample elements include $120^\circ$
rotations (order $3$), $180^\circ$ rotations (order $2$), and $90^\circ$
rotations (order $4$).

\paragraph{The determinant is always $+1$.} Every one of the $24$ holonomies is a
\emph{proper} rotation:
\begin{equation}
  \det R(W) = +1 \quad\text{for all automorphisms } W.
\end{equation}
No closed loop in code space reverses the orientation of the logical Bloch
frame. The reachable inside-twists form a genuine nontrivial rotation group, but
that group is $\mathrm{SO}(3)$-valued: it never touches the orientation-reversing
component of $\mathrm{O}(3)$.

\section{The missing twist is time reversal}

The absence of $\det R = -1$ is not an accident of this code, and we are careful
to say so. Unitary conjugation $\rho \mapsto U\rho U^\dagger$ acts on the Bloch
sphere by proper rotations only; orientation-reversing Bloch-frame maps are
implemented by \emph{antiunitary} operators. By Wigner's theorem~\cite{wigner},
every symmetry of quantum mechanics is either unitary or antiunitary, and time
reversal $T$ is the standard antiunitary symmetry that reverses temporal
orientation. An observer confined to unitary self-operations therefore cannot,
by closed self-transport, acquire the mirror twist:
\begin{equation}
  \boxed{
  \begin{gathered}
  \text{reachable inside-holonomies} = \Sfour \subset \mathrm{SO}(3),\\
  \text{missing sector} = \text{antiunitary, time-reversing twists}.
  \end{gathered}}
\end{equation}
The operation-type reserved for a vantage outside the unitary description is the
one that runs the film backwards. In the program's slogan, temporal orientation
is the currency of self-containment: the inside-operations form a group with no
orientation-reversing element, and the sector they cannot reach is precisely the
arrow's reversal.

\section{Hardware witness}

The \emph{nontrivial} content --- that the reachable holonomies form a real
rotation group rather than collapsing to the identity --- is not guaranteed by
unitarity and is worth measuring. We prepared the $\code$ tile on
\texttt{ibm\_kingston} using the same routed tile-preparation circuit (42
two-qubit gates), then applied the uniform transversal Clifford layer
$K^{\otimes 5}$ realizing an order-$3$ automorphism. The logical frame rotates by
$120^\circ$: the post-loop logical state overlaps the
\emph{rotated} prediction with $+0.898$ and the \emph{unrotated} reference with
$-0.003$ (IBM job \texttt{d958plnu62ks7396rgmg}). The twist is real and it is a
proper rotation; the mirror twist is, as predicted, nowhere on the device.

\section{Relation to prior work, and honest scope}

The statement ``unitary conjugation cannot implement time reversal'' is textbook.
What is, to our knowledge, unstaked is the concrete finite-code, hardware-backed
form of it: that for an explicit self-referential code the reachable
inside-holonomies are exactly $\Sfour$ and the missing orientation-reversing
sector is the antiunitary time-reversal sector. Three operational results frame
the same boundary from different angles and
should be read alongside this note. Manikandan and Jordan~\cite{manikandan}
characterize the time-reversal symmetry of generalized measurements with past and
future boundary conditions. Di Biagio, Don\`a and Rovelli~\cite{dibiagio} locate
the operational arrow of time in assumptions about which quantities are known to
an agent. Chiribella, Aurell and \.Zyczkowski~\cite{chiribella} prove a no-go for
extending the time reversal of unitary dynamics to all quantum evolutions, with
conditioning on past outcomes recovering the standard operational theory. Our
result is complementary and concrete: a physical code, an exhaustive holonomy
census, and a measured twist.

\paragraph{Scope.} We state the limitation plainly. The \emph{exclusion} of the
mirror twist is generic to any unitary framework --- it is Wigner's theorem, not
a property of the $\code$ code --- so this note does not claim a new theorem about
time. Its contribution is threefold and entirely concrete: (i) the reachable
inside-twists are computed exactly and shown to form the specific nontrivial group
$\Sfour$, the full logical Clifford symmetry; (ii) the unreachable sector is
identified as the antiunitary time-reversal sector, tying the census to the
program's temporal-orientation hypothesis; and (iii) a nontrivial element is
realized and measured on hardware. The result is a witness, made physical, of
where the boundary of self-enactable operations lies --- and that the boundary is
drawn at the arrow-reversing sector rather than at ordinary Clifford rotations.

\begin{thebibliography}{9}
\bibitem{laflamme} R.~Laflamme, C.~Miquel, J.~P.~Paz, W.~H.~Zurek,
  \emph{Perfect quantum error correcting code}, Phys. Rev. Lett. \textbf{77},
  198 (1996), \href{https://arxiv.org/abs/quant-ph/9602019}{arXiv:quant-ph/9602019}.
\bibitem{happy} F.~Pastawski, B.~Yoshida, D.~Harlow, J.~Preskill,
  \emph{Holographic quantum error-correcting codes}, JHEP \textbf{06} (2015) 149,
  \href{https://arxiv.org/abs/1503.06237}{arXiv:1503.06237}.
\bibitem{wigner} E.~P.~Wigner, \emph{Group Theory and its Application to the
  Quantum Mechanics of Atomic Spectra} (Academic Press, 1959); antiunitarity of
  time reversal, Ch.~26.
\bibitem{manikandan} S.~K.~Manikandan, A.~N.~Jordan, \emph{Time reversal symmetry
  of generalized quantum measurements with past and future boundary conditions},
  \href{https://arxiv.org/abs/1801.04364}{arXiv:1801.04364}.
\bibitem{dibiagio} A.~Di Biagio, P.~Don\`a, C.~Rovelli, \emph{The arrow of time in
  operational formulations of quantum theory}, Quantum \textbf{5}, 520 (2021),
  \href{https://arxiv.org/abs/2010.05734}{arXiv:2010.05734}.
\bibitem{chiribella} G.~Chiribella, E.~Aurell, K.~\.Zyczkowski, \emph{Symmetries
  of quantum evolutions}, Phys. Rev. Research \textbf{3}, 033028 (2021),
  \href{https://arxiv.org/abs/2101.04962}{arXiv:2101.04962}.
\end{thebibliography}

\vspace{1em}
\noindent\rule{\linewidth}{0.4pt}\\[2pt]
{\footnotesize \emph{Acknowledgements.} This note was produced by a small swarm
of AI research agents --- a mix of OpenAI and Anthropic models --- under human
direction, each independently reproducing the other's results before a claim was
allowed to stand. All code and data are public;
every hardware claim carries an IBM job ID.}

\end{document}
